Evaluating Cloud-Optimized HDF5 for NASA’s ICESat-2 Mission

### 2025-01-16 paper.qmd
modify headings to fit paper structure

#### **Metadata fragmentation**

@@

Fragmentation occurs when this metadata is spread out across multiple non-contiguous chunks within the file. This makes accessing or modifying data inefficient because the HDF library has to read from multiple, scattered locations in the file. Over time, as the dataset grows and evolves, this fragmentation can compound, leading to degraded performance and increased storage overhead. In particular, operations that involve reading or writing metadata, such as opening a file, checking attributes, or modifying variables, can become slower and more resource-intensive.  When cloud-hosted files are accessed over a network using HTTPS, multiple reads increases the time to access data.

#### **Global API Lock**

Because of the historical complexity of operations with the HDF5 format[@The_HDF_Group_Hierarchical_Data_Format], there has been a necessity to make the library thread-safe and similarly to what happens in the Python language, the simplest mechanism to implement this is to have a global API lock. This global lock is not as big of an issue when we read data from local disk but it becomes a major bottleneck when we read data over the network because each read has to be sequential and latency in the cloud is exponentially bigger than local access [@Mozilla-latency-2024] [@scott-2020].

>>>>

##### _Metadata fragmentation_

@@

Fragmentation occurs when this metadata is spread out across multiple non-contiguous chunks within the file. This makes accessing or modifying data inefficient because the HDF library has to read from multiple, scattered locations in the file (@fig-1). Over time, as the dataset grows and evolves, this fragmentation can compound, leading to degraded performance and increased storage overhead. In particular, operations that involve reading or writing metadata, such as opening a file, checking attributes, or modifying variables, can become slower and more resource-intensive.  When cloud-hosted files are accessed over a network using HTTPS, multiple reads increases the time to access data.

##### _Global API Lock_

Because of the historical complexity of operations with the HDF5 format [@The_HDF_Group_Hierarchical_Data_Format], there has been a necessity to make the library thread-safe and similarly to what happens in the Python language, the simplest mechanism to implement this is to have a global API lock. This global lock is not as big of an issue when we read data from local disk but it becomes a major bottleneck when we read data over the network because each read has to be sequential and latency in the cloud is exponentially bigger than local access [@Mozilla-latency-2024] [@scott-2020].
